### 2021-05-10 tables/build/latency.tex
New tables, expanded writing!

\begin{tabular}{@{}ccS[table-format=4.3]S[table-format=4.3]S[table-format=4.3]S[table-format=4.3]S[table-format=4.3]S[table-format=4.3]@{}}
\toprule\multicolumn{1}{c}{Datatype} & \multicolumn{1}{c}{Machine/FW} & \multicolumn{3}{c}{State-Action Latency (\si{\micro\second})} & \multicolumn{3}{c}{State-Update Time (\si{\micro\second})}\\
\cmidrule(lr){3-5}\cmidrule(lr){6-8} & & \multicolumn{1}{c}{Median} & \multicolumn{1}{c}{\nth{99}} & \multicolumn{1}{c}{\num{99.99}\nthscript{th}} & \multicolumn{1}{c}{Median} & \multicolumn{1}{c}{\nth{99}} & \multicolumn{1}{c}{\num{99.99}\nthscript{th}}\\ \midrule
Float & Collector & 515.94 & 606.06 & 725.03 & 606.06 & 636.82 & 833.99\\
 & MidServer & 1069.07 & 1125.1 & 1508.0 & 1260.04 & 1605.99 & 1719.864\\
\cmidrule(lr){1-8}
Int32 & \approachshort{}-\Indfw{} & 185.040 & 185.373 & 185.707 & 230.707 & 231.200 & 231.633\\
 & \approachshort{}-\Coopfw{} & \bfseries 44.133 & \bfseries 44.280 & \bfseries 44.320 & \bfseries 81.287 & \bfseries 81.760 & \bfseries 82.447\\
Int16 & \approachshort{}-\Indfw{} & 195.320 & 195.680 & 195.953 & 242.040 & 242.587 & 243.547\\
 & \approachshort{}-\Coopfw{} & 46.293 & 46.453 & 46.493 & 84.347 & 84.800 & 85.567\\
Int8 & \approachshort{}-\Indfw{} & 193.413 & 193.733 & 194.247 & 240.173 & 240.707 & 241.547\\
 & \approachshort{}-\Coopfw{} & 45.920 & 46.013 & 46.040 & 83.733 & 84.213 & 85.160\\
\bottomrule
\end{tabular}