### 2021-05-21 tables/build/latency-only32.tex
Shorter tables (and stop making Windows cry)

\begin{tabular}{@{}ccS[table-format=4.3]S[table-format=4.3]S[table-format=4.3]S[table-format=4.3]S[table-format=4.3]S[table-format=4.3]@{}}
\toprule\multicolumn{1}{c}{Datatype} & \multicolumn{1}{c}{Machine/FW} & \multicolumn{3}{c}{State-Action Latency (\si{\micro\second})} & \multicolumn{3}{c}{State-Update Time (\si{\micro\second})}\\
\cmidrule(lr){3-5}\cmidrule(lr){6-8} & & \multicolumn{1}{c}{Median} & \multicolumn{1}{c}{\nth{99}} & \multicolumn{1}{c}{\num{99.99}\nthscript{th}} & \multicolumn{1}{c}{Median} & \multicolumn{1}{c}{\nth{99}} & \multicolumn{1}{c}{\num{99.99}\nthscript{th}}\\ \midrule
Float & Collector & 515.94 & 606.06 & 725.03 & 606.06 & 636.82 & 833.99\\
 & MidServer & 1069.07 & 1125.1 & 1508.0 & 1260.04 & 1605.99 & 1719.864\\
Int32 & \approachshort{}-\Indfw{} & 185.040 & 185.373 & 185.707 & 230.707 & 231.200 & 231.633\\
 & \approachshort{}-\Coopfw{} & \bfseries 44.133 & \bfseries 44.280 & \bfseries 44.320 & \bfseries 81.287 & \bfseries 81.760 & \bfseries 82.447\\
\bottomrule
\end{tabular}