### 2021-05-21 tables/build/tput-only32.tex

\begin{tabular}{@{}ccS[table-format=2]S[table-format=3.3]S[table-format=2.3]S[table-format=1.3]S[table-format=1.3]@{}}
\toprule\multicolumn{1}{c}{Datatype} & \multicolumn{1}{c}{Machine/FW} & \multicolumn{1}{c}{Workers} & \multicolumn{2}{c}{Throughput (k actions/s)} & \multicolumn{2}{c}{Throughput/core (k actions/s)}\\
\cmidrule(lr){4-5}\cmidrule(lr){6-7} & & & \multicolumn{1}{c}{Offline} & \multicolumn{1}{c}{Online} & \multicolumn{1}{c}{Offline} & \multicolumn{1}{c}{Online}\\ \midrule
Float & Collector & 4 & 7.673 \pm 0.049 & 1.627 \pm 0.031 & 1.918 \pm 0.012 & \multicolumn{1}{c}{---}\\
 & MidServer & 6 & 5.584 \pm 0.030 & 0.791 \pm 0.012 & 0.931 \pm 0.005 & \multicolumn{1}{c}{---}\\
Int32 & \approachshort{}-\Indfw{} & 32 & \bfseries 172.956 \pm 0.171 & 4.335 \pm 0.005 & \bfseries 5.405 \pm 0.005 & \multicolumn{1}{c}{---}\\
 & \approachshort{}-\Coopfw{} & 32 & 22.670 \pm 0.073 & \bfseries 12.308 \pm 0.043 & 0.708 \pm 0.002 & \bfseries 0.385 \pm 0.001\\
\bottomrule
\end{tabular}