### 2021-12-01 poster/render-table.tex
CoNEXT Poster rendered/done.

\documentclass[%
    ,float=false % this is the new default and can be left away.
    ,preview=true
    % ,class=scrartcl
%    ,fontsize=20pt
    ]{standalone}

   \usepackage{graphicx}

\usepackage{booktabs}
\newcommand{\approach}{On Path Learning}
\newcommand{\approachshort}{OPaL}
\newcommand{\Coopfw}{\emph{CoOp}}
\newcommand{\coopfw}{\Coopfw}
\newcommand{\Indfw}{\emph{Ind}}
\newcommand{\indfw}{\Indfw}
\newcommand{\inring}{\textsc{In}}
\newcommand{\outring}{\textsc{Out}}
\newcommand{\seidr}{Sei\dh{}r}
\newcommand{\seidrsafe}{Seidr}

\usepackage[per-mode=symbol]{siunitx}
\sisetup{detect-all, range-phrase=--, range-units=single, detect-weight=true, table-format=1.3}
\DeclareSIUnit{\packet}{p}

\usepackage[sfdefault]{AlegreyaSans}

\begin{document}

\resizebox{\linewidth}{!}{
\begin{tabular}{@{}ccS[table-format=2]S[table-format=3.3]S[table-format=2.3]S[table-format=1.3]S[table-format=1.3]@{}}
\toprule\multicolumn{1}{c}{Datatype} & \multicolumn{1}{c}{Machine/FW} & \multicolumn{1}{c}{Workers} & \multicolumn{2}{c}{Throughput (k actions/s)} & \multicolumn{2}{c}{Throughput/core (k actions/s)}\\
\cmidrule(lr){4-5}\cmidrule(lr){6-7} & & & \multicolumn{1}{c}{Offline} & \multicolumn{1}{c}{Online} & \multicolumn{1}{c}{Offline} & \multicolumn{1}{c}{Online}\\ \midrule
Float & Collector & 4 & 7.673 \pm 0.049 & 1.627 \pm 0.031 & 1.918 \pm 0.012 & \multicolumn{1}{c}{---}\\
 & MidServer & 6 & 5.584 \pm 0.030 & 0.791 \pm 0.012 & 0.931 \pm 0.005 & \multicolumn{1}{c}{---}\\
Int32 & \approachshort{}-\Indfw{} & 32 & \bfseries 172.875 \pm 0.229 & 4.333 \pm 0.005 & \bfseries 5.402 \pm 0.007 & \multicolumn{1}{c}{---}\\
 & \approachshort{}-\Coopfw{} & 32 & 29.166 \pm 0.173 & \bfseries 16.141 \pm 0.073 & 0.911 \pm 0.005 & \bfseries 0.504 \pm 0.002\\
\bottomrule
\end{tabular}
}

\end{document}